---
title: "R Basics"
author: "Will Ju"
date: "`r Sys.Date()`"
output: 
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
```

# vectors and lists

Vectors and lists are the most basic building structures of `R`

+ vector

```{r}
num_vec <- c(1, 2, 3)

char_vec <- c('apple', 'banana', 'c')

bool_vec <- c(TRUE, FALSE, TRUE)
```

---

+ list

```{r}
will <- list(
  name = 'Will',
  age = 27,
  lives_in_Ames = TRUE
)

will
```

---

# control structures

+ if statements
+ for statements
+ while statements
+ repeat statements
+ break and next statements
+ switch statements

---

# control structures

+ if statements

```{r}
if (will$lives_in_Ames) {
  print("Ames is good")
}
```

+ for statements

```{r}
for (num in num_vec) {
  print(num)
}
```

---

# user-defined functions

```{r}
plus_a_b <- function(arg1, arg2) {
  arg1 + arg2
}

plus_a_b(3, 10)
```

---

# pipe operator

`obj %>% function(arg2)` is equivalent to `function(obj, org2)`

```{r}
3 %>% plus_a_b(10)
```

---

# dataframe

```{r}
data("mtcars")

head(mtcars)
```

---
