---
title: "R Basics"
author: "Will Ju"
date: ""
output: 
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

# Outline

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

- R is a calculator

- vectors and indices

- data as a data.frame object

- extracting pieces

- five commands to look at objects

---

## R is a calculator

- Basic algebra is the same as calculator/mathematics

- explicit operators: `2*x` not `2x`,  `2^p` instead of `2p`

- Applying a function is similar to other programming language

- Making a variable, use `<-` (using `=` is NOT encouraged)

- Everything in R is a vector

---

## Examples


| Math   | Type          | Code           | 
|:--- |:-------------:|-------------:|
|  $$x = \frac{2}{3}$$ | Assignments      | `x <- 2/3` |
| $$\sqrt{x}$$ | Functions  | `sqrt(x)` |
| $$y = \left( \begin{array}{c} 1\\4\\5\\2\end{array}\right)$$ | Vectors  | `y <- c(1, 4, 5, 2) ` |
| $$y_2$$ | Indices | `y[2]`| 

---

## More Examples

| Math   | Type          | Code           | 
|:--- |:-------------:|-------------:|
| $$\sum_{i=1}^{4} y_i$$ | Mathematical Operators | `sum(y)` |
| $$2y$$ |  | `2*y` |

---
class: inverse

## Your Turn (5 min) 


- Introduce vector $x$ defined as $$x = \left( \begin{array}{c} 4\\1\\3\\9\end{array}\right)$$

- Introduce vector $y$ defined as  $$y = \left( \begin{array}{c} 1\\2\\3\\5\end{array}\right)$$


- Calculate the Euclidean distance between the two vectors $x$ and $y$, defined as 

$$d = \sqrt{\sum_{i=1}^4 (x_i - y_i)^2}$$

---

## Vectors and Indices

Vectors are the most basic building structures of `R`

+ there are different types of vectors

```{r}
num_vec <- c(1, 2, 3)

char_vec <- c('apple', 'banana', 'cherry')

bool_vec <- c(TRUE, FALSE, TRUE)
```

+ In `R`, indices start at 1 instead of 0

---

## data as a data.frame object

What is a typical dataset look like?

```{r}
head(mtcars)
```

- each row represents one observation
- each column represents one variable (feature)

---

## data as a data.frame object

How to create a dataset?

- values of a variable is stored in a vector
- vectors are collected by a `data.frame()`

```{r}
mydata <- data.frame(
  num_vec,
  char_vec,
  bool_vec
)
mydata
```

---

## Getting help within R

If you want to know what a specific `command` is doing:

```{r, eval=FALSE}
?command

help("command") 

help.search("command")

??command
```


---

## Loading class data

- Some R packages have in-built datasets

- For this class, there is an R package available on github 

- we will use a function `install_github` from `remotes` package. So if you haven't installed `remotes` package, please do so first

```{r eval=FALSE}
install.packages("remotes")
```

- Installing/Updating `classdata` package (once every so often):
```{r eval=FALSE}
remotes::install_github("heike/classdata")
```

- Make the data available (every time you start R):
```{r}
library(classdata)
```


---
class: inverse
## Your Turn (5 min) 


- Install the package `classdata` on your machine <br>

- Make the package active in your current R session:
```{r, eval=FALSE}
library(classdata)
```

- Check the R help on the dataset `fbi`<br>

- What happens if you just type in the name of the dataset?


---

## Extracting parts of objects

- use indices

```
x[row_index, col_index]
x[1:5, 2:3]
x[c(1,5,6), c(3,4,5)]
```

- For a 2D object (a typical datasets) with column (variable) names

```
x$variable
x[, var_names]
x[, c("State", "Year")]
```

<font color="darkblue">Try these commands out for yourself on the `fbi` data.</font>

---

## five commands to look at objects

for object `x`, we can try out the following commands:

- `x`

- `head(x)`

- `summary(x)`

- `str(x)`

- `dim(x)`

<br><br><br>

<font color="darkblue">Try these commands out for yourself on the `fbi` data.</font>

---

## `str` stands for *structure*

- `str` shows us the **str**ucture of an object

- fbi is a data frame with columns (variables) and rows (records)

```{r}
str(fbi)
```

---

## Statistical summaries

Elements of the five point summary: <br>
`mean, median, min, max, quartiles`

Other summary statistics:<br>
`range, sd, var`

Summaries of two variables:<br>
`cor, cov`

---
class: inverse
## Your turn 

- Look at the first 10 data records of the `fbi` data

- Compute mean and standard deviation for the number of counts. Why do you get NAs? (read `?NA`)

- Advanced:  Read `?mean` and `?sd`, and fix missing value problem

---

## More about R

- lists

- control structures

- user-defined functions

---

## lists

`list` is also a fundamental building structure of `R`. Vector stores values of the same type, but `list` can store values of different types.

```{r}
will <- list(
  name = 'Will',
  age = 28,
  lives_in_Ames = TRUE
)

will
```

---

# control structures

+ if statements
+ for statements
+ while statements
+ repeat statements
+ break and next statements
+ switch statements


---

# control structures

+ if statements

```{r}
if (will$lives_in_Ames) {
  print("Ames is good")
}
```

+ for statements

```{r}
for (num in num_vec) {
  print(num)
}
```

---

# user-defined functions

```{r}
plus_a_b <- function(arg1, arg2) {
  arg1 + arg2
}

plus_a_b(3, 10)
```


